Probe or Wait: Handling tail losses using Multipath TCP

### 2017-04-27 main.tex
Updates to sec. IV

\begin{abstract}
Packet losses are known to affect the performance of latency sensitive Internet applications such as media streaming and gaming. 
Transport protocols recover from packet loss in order to provide reliable end-to-end communication and improve user experience. The efficiency of 
loss recovery mechanisms influences the completion time of flows, and thus also the application performance as perceived by the
end user. In this paper we focus on state-of-the-art loss recovery mechanisms for TCP 
and Multipath TCP. We use controlled tail loss scenarios to evaluate the performance of loss recovery mechanisms, and based on the observations, 
we propose an enhanced tail loss recovery mechanism for Multipath TCP, to improve the loss recovery time. Our experiment results, using the Linux 
Multipath TCP implementation, show consistent end-to-end latency performance improvement in considered scenarios. 
\end{abstract}

@@

\section{Observations and Discussion}\label{disc}
This section provides the performance analysis of TCP and MPTCP for the considered cases of single packet tail loss, two packet tail loss and 
retransmission or probe loss.
\subsection{TCP}
Retransmission behavior of TCP provides a base case for MPTCP performance as each individual MPTCP subflow is a TCP flow in itself. Our analysis
starts with the results using TCP and compares it with that of MPTCP. Figure~\ref{tcpf} represents the comparison of TCP burst completion times. Path
asymmetry is irrelevant in this case as TCP uses a single path.

In the case of single packet tail loss, as shown in Figure~\ref{t1p}, there is no difference in burst completion between RTO and TLP. The 
retransmission timeout and the probe timeout are the same as there is only one outstanding packet. This is a special case in the TLP
specification that adds 200\,ms to the PTO accommodating the case of delayed ACKs. The TLP implementation in Linux accounts for the delayed 
ACK and waits 200\,ms before sending a probe. The minimum of PTO and RTO is considered for PTO. In the case of two packet tail loss, as shown 
in Figure~\ref{t2p}, TLP performs better than the RTO retransmission for lower delay values. As the delay on the loss path increases RTO retransmission performs 
slightly better than TLP. In 120\,ms case, PTO is very close to RTO expiration and the second packet is recovered using ER which incurs additional 
delay in case of TLP. In RTO retransmission, all unacknowledged packets are retransmitted on a RTO expiration.

Hence retransmission timeout is the mode of loss recovery in both RTO and TLP. This can be avoided by tuning the ACKs as discussed 
in~\cite{Rajiullah:2015}. In the case of two packet tail loss, TLP performs better than RTO retransmission as it employs a combination of TLP 
and ER to recover from losses. If the probe or retransmission is lost, then TLP still performs better than RTO retransmission as it waits for one 
probe timeout and one RTO experiation instead of two RTO expirations, thus avoiding the exponential backoff of the RTO.

@@

MPTCP uses an RTT based scheduler by default, that uses the lowest RTT path among the available paths to transmit the data until its congestion 
window is full. Subsequently, the next lowest RTT path is chosen. MPTCP performance is sensitive to asymmetry in paths. The burst completion times 
on single packet tail loss calculated for each delay configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no 
difference in performance between RTO retransmission and TLP for cases when both paths are symmetric or the first path has lowest RTT. RTO performs 
better than TLP when the first path is not the lowest RTT path. In such scenarios~(30\,ms-20\,ms,120\,ms-20\,ms), the RTO retransmission happens 
in the second path with lowest RTT following the behavior depicted in Figure~\ref{timing1P}.

@@

In the case of single packet tail loss, the last outstanding packet is sent as probe and in case of n-degree tail loss, the latest packet is sent 
as probe to trigger fast retransmit of the packets in between. In order to observe the performance of n-degree tail loss, we consider dropping the 
last two packets. The results of two packet tail loss are shown in Figure~\ref{2p}. In this case, TLP performs better than the other 
two settings in all scenarios except when packets are lost on a high delay interface. In this scenario, the MPTCP retransmission follows the 
behavior depicted in Figure~\ref{timing2P} and the delay difference is large enough that the retransmission of the last packet on the alternate 
path is faster than the probe on the primary path.

To further study the performance of TLP in the case when there is longer path interruption on one path, we tried to drop the probe packet along 
with the last packet. In this scenario, the RTO setting results in much lower burst completion times. Comparing with the performance of TLP in 
TCP from Figure~\ref{t1pp}, where we saw that use of TLP improved the burst completion time, TLP in MPTCP did not improve the performance. The 
reason lies in the way the retransmission and loss probe transmission is carried out in MPTCP. Retransmission occurs after an RTO time on both 
paths. Even if the retransmission fails on one path, the client receives the packet on the other path as shown in Figure~\ref{timing1PP}. For TLP, 
in the current implementation, the loss probe is sent on the actual path of the first transmission but not on both paths. This incomplete impartation 
of TLP from TCP to MPTCP led to the increase in burst completion time by an RTO for MPTCP.

>>>>

\begin{abstract}
Packet losses are known to affect the performance of latency sensitive applications in the Internet such as media streaming and gaming. 
Transport protocols recover from packet loss in order to provide reliable end-to-end communication and improve user experience. The efficiency of 
loss recovery mechanisms influences the completion time of flows, and thus also the application performance as perceived by the
end user. In this paper we focus on state-of-the-art loss recovery mechanisms for TCP 
and Multipath TCP. We use controlled tail loss scenarios to evaluate the performance of loss recovery mechanisms, and based on the observations, 
we propose an enhanced tail loss recovery mechanism for Multipath TCP, to improve the loss recovery time. Our experiment results using the Linux 
Multipath TCP implementation show consistent end-to-end latency performance improvement in considered scenarios. 
\end{abstract}

@@

\section{Observations and Discussion}\label{disc}
This section provides the performance analysis of TCP and MPTCP for the considered cases of single packet tail loss, two packet tail loss and 
retransmission or probe loss.
\subsection{TCP}
Retransmission behavior of TCP provides a base case for MPTCP performance as each individual MPTCP subflow is a TCP flow in itself. Our analysis
starts with the results using TCP and compares it with that of MPTCP. Figure~\ref{tcpf} represents the comparison of TCP burst completion times
in the considered scenarios. Path
asymmetry is irrelevant in this case as TCP uses a single path.

In the case of single packet tail loss, as shown in Figure~\ref{t1p}, there is no difference in burst completion between RTO and TLP. The 
retransmission timeout and the probe timeout are the same as there is only one outstanding packet. This is a special case in the TLP
specification that adds 200\,ms to the PTO accommodating the case of delayed ACKs. The TLP implementation in Linux accounts for the delayed 
ACK and waits 200\,ms before sending a probe. The minimum of PTO and RTO is considered for the PTO. Hence in this special case the value of 
the retransmission timeout is used in the loss recovery in both RTO and TLP. This can be avoided by tuning the ACKs as discussed 
in~\cite{Rajiullah:2015}.

In the case of two packet tail loss, as shown 
in Figure~\ref{t2p}, TLP performs better than the RTO retransmission for lower delay values. As the delay on the loss path increases, RTO retransmission performs 
slightly better than TLP. In the 120\,ms case, PTO is very close to RTO expiration and in the case of TLP the second packet is recovered 
using ER which incurs additional delay. In RTO retransmission, the second packet is retransmitted immediately when the acknowledgement
for the RTO retransmission is received.

If the probe or retransmission is lost, as shown 
in Figure~\ref{tcpf}, then TLP performs better than RTO retransmission as it waits for one 
probe timeout and one RTO experiation instead of two RTO expirations, thus avoiding the exponential backoff of the RTO.

@@

MPTCP uses an RTT based scheduler by default, that uses the lowest RTT path among the available paths to transmit the data until its congestion 
window is full. Subsequently, the next lowest RTT path is chosen. MPTCP performance is sensitive to asymmetry in paths. The burst completion times 
on single packet tail loss calculated for each delay configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no 
difference in performance between RTO retransmission and TLP for cases when both paths are symmetric or the first path has the lowest RTT. RTO performs 
better than TLP when the first path is not the lowest RTT path. In such scenarios~(30\,ms-20\,ms,120\,ms-20\,ms), the RTO retransmission happens 
in the second path with lowest RTT following the behavior depicted in Figure~\ref{timing1P}.

@@

In the case of single packet tail loss, the last outstanding packet is sent as probe and in case of n-degree tail loss, the latest packet is sent 
as probe to trigger fast retransmit of the packets in between. In order to observe the performance of n-degree tail loss, we consider dropping the 
last two packets. The results of two packet tail loss are shown in Figure~\ref{2p}. In this case, TLP performs better than the RTO 
setting in all scenarios except when packets are lost on a high delay interface. In this scenario, the MPTCP retransmission follows the 
behavior depicted in Figure~\ref{timing2P} and the delay difference is large enough that the retransmission of the last packet on the alternate 
path is faster than the probe on the primary path.

To further study the performance of TLP in the case when there is a longer path interruption on one path, we tried to drop the probe packet along 
with the last packet. In this scenario, the RTO setting results in much lower burst completion times. Comparing with the performance of TLP in 
TCP from Figure~\ref{t1pp}, where we saw that the use of TLP improved the burst completion time, TLP in MPTCP did not improve the performance. The 
reason lies in the way the retransmission and loss probe transmission are carried out in MPTCP. After an RTO, retransmissions occur on both 
paths. Even if the retransmission fails on one path, the client receives the packet on the other path as shown in Figure~\ref{timing1PP}. For TLP, 
in the current implementation, the loss probe is sent on the actual path of the first transmission but not on both paths. This incomplete impartation 
of TLP from TCP to MPTCP led to the increase in burst completion time by an RTO for MPTCP.
